### 2014-07-10 paper.tex
Added LaTeX version of paper.

\documentclass[11pt, oneside]{amsart}

\usepackage{amsmath}
\usepackage{amssymb}

%\usepackage{multibbl}
%\usepackage{bibtopic}

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
\usepackage[latin9]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}

%\setcounter{secnumdepth}{3}
%\setcounter{tocdepth}{3}

\usepackage[bookmarks, bookmarksopen, bookmarksnumbered]{hyperref}
\usepackage[all]{hypcap}
\urlstyle{rm}

\definecolor{orange}{rgb}{1.0,0.3,0.0}
\definecolor{violet}{rgb}{0.75,0,1}
\definecolor{darkgreen}{rgb}{0,0.6,0}
\definecolor{cyan}{rgb}{0.2,0.7,0.7}
\definecolor{blueish}{rgb}{0.2,0.2,0.8}

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\huttonnote}[1]{ {\textcolor{magenta}    { ***Eric:      #1 }}}

% Don't use tt font for urls
\urlstyle{rm}

\begin{document}

\title[]{Building Sustainable Software - The CSDMS-2.0 Approach}

\author{Eric W. H. Hutton$^{\dag}$, Mark D. Piper$^{\dag}$, Irina Overeem$^{\dag}$, Albert J. Kettner$^{\dag}$}
\thanks{{}$^{\dag}$ Community Surface Dynamics Modeling System, Boulder, CO}

\begin{abstract}

Our abstract goes here.

\end{abstract}

\maketitle

\section{The Community Surface Dynamics Modeling System}

The Community Surface Dynamics Modeling System (CSDMS)
~\cite{peckham2012component}
is an NSF-funded project that began in 2007. Its mission is to help a diverse
community of surface dynamics modelers and model users work together toward
common goals and standards. Part of this effort involves creating a repository
of open-source models. Another part involves converting models to plug-and-play
components that can be reused via dynamic linking within the CSDMS framework to
create new models. In building its modeling framework, CSDMS has leveraged
several existing, well-established and open-source software tools. For example,
CSDMS uses three tools from the Common Component Architecture (CCA) toolchain:
Babel, Bocca and Ccaffeine. Babel provides interoperability between components
written in different languages; it currently supports C, C++, Fortran, Java,
and Python. Bocca helps with creating CCA-compliant components and managing CCA
component projects. Ccaffeine is a CCA-compliant framework for linking
components into new applications. All three tools were designed to enable
component-based modeling in an HPC environment, but they can also be used on
desktop systems. CSDMS has developed innovative model/component interfaces for
use with the CCA tools, including the Basic Model Interface (BMI), which uses
the CSDMS Standard Names (both described below) and the framework-level
Component Model Interface (CMI). CSDMS has also developed a web-based client
for its modeling framework (the CSDMS Web Modeling Tool, WMT, described below)
that allows users to compose new models by connecting and configuring
components through a simple, browser-based graphical interface.

In order to realize the full power and benefits of component-based modeling, a
modeling framework like CSDMS needs an efficient mechanism to convert as many
open-source models as possible to reusable plug-and-play components. Since this
necessarily requires some involvement from each model's developer, this
mechanism must be designed to:

\begin{itemize}

\item require minimal effort from the model developer,

\item allow the model to continue to be used in a stand-alone manner,

\item not introduce new dependencies into the model,

\item not interfere with the developer's design,

\item not require any modeling framework-specific knowledge on the part of the
      developer,

\item not require the addition of new code which accommodates the needs of
      other models.

\end{itemize}

These requirements became clear during the first few years of the CSDMS project
by working directly with model developers and listening to their concerns and
complaints about early designs and other frameworks.

\section{The Basic Modeling Interface}

The \emph{Basic Modeling Interface} (BMI) is a component-coupling library interface
specification designed by CSDMS~\cite{peckham2012component, syvitski2014plug}.
In this context, a component is software that models a particular environmental
process and can be "plugged into" another component. BMI addresses the
interaction of components that operate within a shared memory space by
providing entry points for: component control (initialize, run, and finalize
steps), data access, and metadata. This allows developers to write applications
that connect components to one another to form new models.

Most models require input variables and produce output variables. In a
component-based modeling framework, like CSDMS, a set of components becomes a
complete model when every component is able to obtain the input variables it
needs from another component. The CSDMS Standard Names were developed to
provide a practical solution to this semantic mediation problem
~\cite{peckham2012component, syvitski2014plug}.
While the CF Convention Standard Names, which were
introduced in the domain of the ocean and atmosphere model domains, have
somewhat overlapping goals, the CSDMS Standard Names provide a more
comprehensive set of naming rules and patterns for creating unique labels for
model variables that are not specific to any particular modeling domain.

CSDMS has found that BMI is an acceptable target for model developers. The use
of BMI has also dramatically reduced the effort required by CSDMS staff to
create and maintain components. There is now really just one universal CMI
wrapper to maintain that BMI-enabled models "plug into". The introduction of the
CSDMS Standard Names provides a practical and elegant solution to the semantic
mediation problem and Babel solves the language interoperability problem.
CSDMS has also assembled a large repository of surface process models that now
includes over 171 open-source models and 56 tools. The repository contains
terrestrial, coastal, marine, hydrological, carbonate and atmospheric models.
Geodynamic, atmospheric and ecosystem models will be added over the next few
years. So far, 55 of the 171 models have been wrapped for use as reusable
plug-and-play components (although only 6 are currently visible in the new WMT).
CSDMS has a steering committee, an executive committee, five working groups, six
focus research groups, an industrial consortium, and over 1150 members as of
June 2014. Members share information through a wiki.

The CSDMS Standard Names are a lingua franca for variable names between model
components. They play an important role in the BMI as they provide a mapping of
a model's internal variable names to a common language used by the BMI getter
and setter functions.

\section{CSDMS Standard Names}

The CSDMS Standard Names are currently being extended to solid Earth modeling
domains, including geodynamics, seismology, electromagnetic induction and
petrology, in the framework of the "Earth System Bridge" EarthCube project led
by PI Peckham. They also serve as the base for the metadata standard for
numerical models (MMF) which is currently in development within CSDMS.

We are acutely aware of multiple other efforts to provide cross-domain
ontologies in the realm of geosciences. PIs Peckham and Kelbert are involved
with the initial steps in semantic crosswalk development recently initialized
within the EarthCube community that aim to achieve synergy between these
currently divergent metadata ontologies.

\section{Reusable Components}

\subsection{The CSDMS Modeling Framework}

\subsection{Connecting Components with the Web Modeling Tool}

The CSDMS Web Modeling Tool (WMT; https://csdms.colorado.edu/wmt) is the
web-based successor to the desktop Component Modeling Tool. WMT is a web
application that provides an Ajax client-side graphical interface (the WMT
client) and a RESTful server-side database and API (the WMT server) that allows
users to build and run coupled Earth system models on a high-performance
computing cluster (HPCC) from a web browser.
WMT was designed with four objectives:
\begin{itemize}

\item  \emph{Accessibility}. As a web-based application, if you have access
to the Internet, you have access to WMT.

\item  \emph{Integration}. Easily hyperlink from WMT to resources on the CSDMS
portal - including model documentation, labs, lectures, tutorials and
movies - or to other resources on the Internet.

\item \emph{Portability}. WMT has a native JavaScript interface, so it can be
accessed on any modern web browser, including tablet and mobile
versions of browsers.

\item \emph{Maintenance}. Because modern browsers tend to adhere to web
standards, which lead to fewer cross-compatibility issues than
operating systems, only one version of WMT needs to be developed
and maintained.

\end{itemize}

With WMT, a user can:

\begin{itemize}

\item select a Common Component Architecture (CCA) component model
      from a list to run in standalone mode;

\item build a coupled model from multiple CCA components organized as nodes
      of a tree structure;

\item  view and edit the parameters of the model components;

\item save models to a server, where they can be accessed on any computer
      connected to the Internet;

\item share saved models with others in the community;

\item run a model by connecting to a remote HPCC where the components are
      installed.

\end{itemize}

Although WMT is web-based, the building and configuring of a model can be done
offline. Reconnection is necessary only when saving a model and submitting it
for a run.

\section{Digital Object Identifiers for Numerical Models}

CSDMS encourages and supports
the effort of developers to provide their numerical models as truly open source
~\cite{ince2012case}.
(Ince et al., 2012). All code in CSDMS is open source. Source code exposes the
scientific hypotheses embodied in a numerical model. Solutions to a set of
equations can take numerous forms, each with their pyramid of assumptions and
limitations. Code transparency allows for full peer review and replication of
results - the foundation of modern science. Code transparency also allows for
reuse in new and clever ways, and reduces redundancy. At the same time CSDMS
strives to ensure that model developers receive recognition for their work, even
when code is submitted which is not (yet) described in a scientific journal.
Therefore CSDMS has adopted the Digital Object Identifier (DOI). The DOI system
provides a unique identification to content that is available on digital
networks~\cite{paskin2010}. The DOI system was initiated in 1997 and contained
mainly DOIs to publishing applications. Since 2005 DOIs were also made available
for research data~\cite{paskin2005digital}. CSDMS appears to be the first to
assign Digital Object Identifiers (DOIs) to numerical source code, as of 2013.
The advantages of adopting a DOI system for models are:

\begin{itemize}

\item Guarantee credit to a model developer.

\item Easily reuse and replicate research as other researchers can directly
      locate a referenced model.

\item Providing a higher visibility, as content with a DOI is 5 times more
      likely to deliver active links than content without.

\item Giving the opportunity to for example funding agencies to track usage,
      so to measure impact.

\end{itemize}

CSDMS collaborates with Integrated Earth Data Applications
(IEDA), a formal Publication Agent of the DOI system through the German National
Library of Science and Technology, to assign unique identifiers only for those
models that contain metadata and are physically part of the CSDMS repository. An
archive of all numerical models of the CSDMS model repository that have a DOI
together with limited metadata as well as the source code for each model is
provided to IEDA as a backup to guaranty access to model information and its
source code beyond the CSDMS program; a DOI for an object is permanent, whereas
its location and other metadata may change in future.

A new DOI is provided for each new version of a model (i.e. major upgrade of the
source code involving a change in version number). CSDMS uses Apache Subversion,
better known as SVN, for tracking source code versioning and revision control so
that current and past releases and changes can be accessed through the web. As
of March 2014, 109 models within the CSDMS model repository have a DOI. Model
source code can be viewed as "data" and therefore CSDMS endorses citations
defined by DataCite guidelines~\cite{brase2009datacite}. Following these guidelines, CSDMS
strongly recommends the following structure for citing a model:
\emph{ModelDeveloper} (\emph{PublicationYear}). \emph{ModelName}.
\emph{ModelVersion}. \emph{Identifier}.

\section{Acknowledgements}

The CSDMS Integration Facility operates under continuing grant 0621695 from the
US National Science Foundation.

\bibliographystyle{plain}

\bibliography{wssspe-citations}{}

\end{document}

### 2014-07-11 paper.tex
Added a figure of WMT.
Added everyone's stuff.

@@
 \usepackage{graphicx}
+%\usepackage{graphics}
 \usepackage[latin9]{inputenc}
@@
 \usepackage{wrapfig}
+\usepackage{framed,color}
+\usepackage{fancybox}
 
@@
 
+\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
+
 \newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
 \newcommand{\huttonnote}[1]{ {\textcolor{magenta}    { ***Eric:      #1 }}}
+\DeclareRobustCommand{\csdms}{\textsc{csdms}}
 
@@
 
-\author{Eric W. H. Hutton$^{\dag}$, Mark D. Piper$^{\dag}$, Irina Overeem$^{\dag}$, Albert J. Kettner$^{\dag}$}
+\author{
+  Eric W. H. Hutton$^{\dag}$,
+  Mark D. Piper$^{\dag}$,
+  Albert J. Kettner$^{\dag}$,
+  Irina Overeem$^{\dag}$}
+
 \thanks{{}$^{\dag}$ Community Surface Dynamics Modeling System, Boulder, CO}
@@
 
-Our abstract goes here.
+The Community Surface Dynamics Modeling System (\csdms{}) is an NSF funded
+project whose focus is to aid a diverse community of earth and ocean system
+model \emph{users} and \emph{developers} to use and create robust software
+quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
+earth-system models and tools to an international (67 country) community
+with the goal of building a complete set of tools necessary to model the
+earth system. \csdms is a place modelers look to for access to hundreds open
+source surface-dynamics models and tools, as well as model metadata. Such a
+model repository increases model transparency and helps to eliminate the
+duplication of effors by presenting the current state of modeling efforts.
+To increase software sustainability, composability and interoperability \csdms
+promotes standards that define common modeling interfaces, semantic mediation
+between models, and model metadata. Through online resources and workshops,
+\csdms promotes software engineering best practices, which are foreign to many
+developers within our modeling community. For example, version control, unit
+testing, continious integration, test-driven development, well-written
+\emph{clean code}.
 
@@
 
-The Community Surface Dynamics Modeling System (CSDMS)
+The Community Surface Dynamics Modeling System (\csdms{})
 ~\cite{peckham2012component}
 is an NSF-funded project that began in 2007. Its mission is to help a diverse
-community of surface dynamics modelers and model users work together toward
+community of surface dynamics model \emph{developers} and model \emph{users}
+work together toward
 common goals and standards. Part of this effort involves creating a repository
 of open-source models. Another part involves converting models to plug-and-play
-components that can be reused via dynamic linking within the CSDMS framework to
-create new models. In building its modeling framework, CSDMS has leveraged
+components that can be reused via dynamic linking within the \csdms framework to
+create new models. In building its modeling framework, \csdms has leveraged
 several existing, well-established and open-source software tools. For example,
-CSDMS uses three tools from the Common Component Architecture (CCA) toolchain:
+\csdms uses three tools from the Common Component Architecture (CCA) toolchain:
 Babel, Bocca and Ccaffeine. Babel provides interoperability between components
@@
 component-based modeling in an HPC environment, but they can also be used on
-desktop systems. CSDMS has developed innovative model/component interfaces for
+desktop systems. \csdms has developed innovative model/component interfaces for
 use with the CCA tools, including the Basic Model Interface (BMI), which uses
-the CSDMS Standard Names (both described below) and the framework-level
-Component Model Interface (CMI). CSDMS has also developed a web-based client
-for its modeling framework (the CSDMS Web Modeling Tool, WMT, described below)
+the \csdms Standard Names (both described below) and the framework-level
+Component Model Interface (CMI). \csdms has also developed a web-based client
+for its modeling framework (the \csdms Web Modeling Tool, WMT, described below)
 that allows users to compose new models by connecting and configuring
@@
 In order to realize the full power and benefits of component-based modeling, a
-modeling framework like CSDMS needs an efficient mechanism to convert as many
+modeling framework like \csdms needs an efficient mechanism to convert as many
 open-source models as possible to reusable plug-and-play components. Since this
@@
 
-These requirements became clear during the first few years of the CSDMS project
+These requirements became clear during the first few years of the \csdms project
 by working directly with model developers and listening to their concerns and
@@
 
+\section{Building a User Community}
+
+\csdms{} provides cyber-infrastructure to promote the quantitative modeling of
+earth surface processes. The US National Science Foundation funds CSDMS, but
+the community is international and includes industry and federal agency
+representatives in addition to members from academic institutions. The
+initiative has been growing with >150 people per year. CSDMS now fosters a
+community of ~1150 scientists who work on prediction of the movement of fluids,
+and the production, erosion, transport, and deposition of sediment and
+nutrients in landscapes and seascapes.
+
+\section{Model Repository}
+
+Software accessibility is a key value of the CSDMS mission.
+CSDMS ensures that code developed by individual researchers or small research teams remains accessible beyond the lifetime of a specific research project.
+Code developers are coaxed into providing their codes under open-source licenses.
+
+CSDMS has built an online model repository that importantly documents the submitted code with metadata, smoke-tests codes on the CSDMS High Performance Computing Cluster, maintains a snapshot of submitted source code in a version-control system, and provides easy downloads of the original source code.
+This repository now contains 172 codes and 56 tools, or 4,074,236 lines of source code (a source code line is defined as any line that is not part of a comment, and is not blank).
+
+We do know our community values this resource, we track downloads of each individual code and the total number of downloads since November 2008 is >13,218 (as of July 2014).
+
+\section{Software sustainability through expanded functionality}
+
+Another important goal of CSDMS is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales.
+Interdisciplinarity and coupling is a critical step towards advancing predictive modeling~\cite{voinov2010community}.
+
+CSDMS develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.
+
+The modeling coupling framework is designed to catalyze Earth-surface process research by empowering a broad community of scientists, students and applied modelers with computing tools, streamlining the process of idea generation and hypothesis testing through linked surface dynamics models, and enabling rapid creation and coupling of models tailored to specific settings, scientific problems, and time scales
+
+\section{CSDMS has an educational mission}
+
+WMT provides the ability to run models with an easy to use graphical interface and intends to serve novice and more experienced modelers alike.
+The WMT beta version was released in May 2014  as a sequel to the CMT.
+IF staff habitually evaluates the effectiveness of this tool and uses the feedback to improve the functionality: modeling courses at the University of Colorado, clinics at the International Summer Institute at the University of Minnesota, and clinics at annual CSDMS Annual Meetings are the dominant forums for this testing.
+
+Over the last 4 years >150 graduate students were enrolled in intensive modeling courses, originating from geography, hydrology, geology, oceanography, earth sciences and engineering departments.
+These courses provided an excellent use case for us to test usability and access to the tool.
+Students reported in post-course surveys that they were unfamiliar with the use of models on a high performance computing cluster beforehand (rated their familiarity as 1.5 average on a scale of 1-5), but ended the courses with familiarity and being more comfortable with running and changing simulations as well as visualizing results (average rating approaches 4 on a Likert scale of 1-5).
+
+Focused hands-on clinics at meetings also allowed more than 150 CSDMS members to experience the ease of use of CMT, and now WMT, and try several examples of how to use the tool.
+
+Course and clinic materials are posted online as educational resources.
+IF staff are developing WMT-specific help pages for all models currently available as components, so that users can easily learn about model parameters and input files.
+In addition, IF staff members are developing web-based video tutorials that show how to run models.
+
 \section{The Basic Modeling Interface}
 
-The \emph{Basic Modeling Interface} (BMI) is a component-coupling library interface
-specification designed by CSDMS~\cite{peckham2012component, syvitski2014plug}.
-In this context, a component is software that models a particular environmental
-process and can be "plugged into" another component. BMI addresses the
-interaction of components that operate within a shared memory space by
-providing entry points for: component control (initialize, run, and finalize
-steps), data access, and metadata. This allows developers to write applications
-that connect components to one another to form new models.
+The \emph{Basic Modeling Interface} (BMI) is a component-coupling library
+interface specification designed by \csdms~\cite{peckham2012component,
+syvitski2014plug}.  In this context, a component is software that models a
+particular environmental process and can be "plugged into" another component.
+
+The BMI specification was designed without any particular model-coupling
+framework in mind.  Rather, it was designed to be framework agnostic. We
+recognize model-coupling frameworks and tools may come and go but the
+functionality a framework requires from its components will be significantly
+more long-lived. Thus, a component should outlive any framework that it might
+operate within.
+
+The BMI identifies the entry points into software components that provides a
+calling appication with the necessary level of control to connect components.
+\csdms, as well as other modeling frameworks such as
+ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
+OMS~\cite{david2002object}, identifies an interface that, at a minimum,
+provides functionality to \emph{initialize}, \emph{update}, and
+\emph{finalize} a component. However, for more tight coupling the components
+must implement an interface to provide data access, and model metadata.
+
+\csdms has found that BMI is an acceptable target for model developers. The use
+of BMI has also dramatically reduced the effort required by \csdms staff to
+create and maintain components. So long as a model component strictly exposes
+a BMI, it can automatically be incorporated into the \csdms model-coupling
+framework where it can then connect to compatible components to form larger
+models.
+
+%The introduction of the
+%\csdms Standard Names provides a practical and elegant solution to the semantic
+%mediation problem and Babel solves the language interoperability problem.
+%\csdms has also assembled a large repository of surface process models that now
+%includes over 171 open-source models and 56 tools. The repository contains
+%terrestrial, coastal, marine, hydrological, carbonate and atmospheric models.
+%Geodynamic, atmospheric and ecosystem models will be added over the next few
+%years.
+
+%So far, 55 of the 171 models have been wrapped for use as reusable
+%plug-and-play components (although only 6 are currently visible in the new WMT).
+
+\section{CSDMS Standard Names}
+
+\huttonnote{Taken from the SSI proposal. I need to rewrite this.}
+
+The \csdms Standard Names are a lingua franca for variable names between model
+components. They play an important role in the BMI as they provide a mapping of
+a model's internal variable names to a common language used by the BMI getter
+and setter functions.
 
 Most models require input variables and produce output variables. In a
-component-based modeling framework, like CSDMS, a set of components becomes a
+component-based modeling framework, like \csdms, a set of components becomes a
 complete model when every component is able to obtain the input variables it
-needs from another component. The CSDMS Standard Names were developed to
+needs from another component. The \csdms Standard Names were developed to
 provide a practical solution to this semantic mediation problem
@@
 introduced in the domain of the ocean and atmosphere model domains, have
-somewhat overlapping goals, the CSDMS Standard Names provide a more
+somewhat overlapping goals, the \csdms Standard Names provide a more
 comprehensive set of naming rules and patterns for creating unique labels for
@@
 
-CSDMS has found that BMI is an acceptable target for model developers. The use
-of BMI has also dramatically reduced the effort required by CSDMS staff to
-create and maintain components. There is now really just one universal CMI
-wrapper to maintain that BMI-enabled models "plug into". The introduction of the
-CSDMS Standard Names provides a practical and elegant solution to the semantic
-mediation problem and Babel solves the language interoperability problem.
-CSDMS has also assembled a large repository of surface process models that now
-includes over 171 open-source models and 56 tools. The repository contains
-terrestrial, coastal, marine, hydrological, carbonate and atmospheric models.
-Geodynamic, atmospheric and ecosystem models will be added over the next few
-years. So far, 55 of the 171 models have been wrapped for use as reusable
-plug-and-play components (although only 6 are currently visible in the new WMT).
-CSDMS has a steering committee, an executive committee, five working groups, six
-focus research groups, an industrial consortium, and over 1150 members as of
-June 2014. Members share information through a wiki.
+%The \csdms Standard Names are currently being extended to solid Earth modeling
+%domains, including geodynamics, seismology, electromagnetic induction and
+%petrology, in the framework of the "Earth System Bridge" EarthCube project led
+%by PI Peckham. They also serve as the base for the metadata standard for
+%numerical models (MMF) which is currently in development within \csdms.
 
-The CSDMS Standard Names are a lingua franca for variable names between model
-components. They play an important role in the BMI as they provide a mapping of
-a model's internal variable names to a common language used by the BMI getter
-and setter functions.
-
-\section{CSDMS Standard Names}
-
-The CSDMS Standard Names are currently being extended to solid Earth modeling
-domains, including geodynamics, seismology, electromagnetic induction and
-petrology, in the framework of the "Earth System Bridge" EarthCube project led
-by PI Peckham. They also serve as the base for the metadata standard for
-numerical models (MMF) which is currently in development within CSDMS.
-
-We are acutely aware of multiple other efforts to provide cross-domain
-ontologies in the realm of geosciences. PIs Peckham and Kelbert are involved
-with the initial steps in semantic crosswalk development recently initialized
-within the EarthCube community that aim to achieve synergy between these
-currently divergent metadata ontologies.
+%We are acutely aware of multiple other efforts to provide cross-domain
+%ontologies in the realm of geosciences. PIs Peckham and Kelbert are involved
+%with the initial steps in semantic crosswalk development recently initialized
+%within the EarthCube community that aim to achieve synergy between these
+%currently divergent metadata ontologies.
 
@@
 
-\subsection{The CSDMS Modeling Framework}
+\subsection{The \csdms Modeling Framework}
 
@@
 
-The CSDMS Web Modeling Tool (WMT; https://csdms.colorado.edu/wmt) is the
+\huttonnote{Taken from the SSI proposal. I need to rework this a bit.}
+
+The \csdms Web Modeling Tool (WMT; https://csdms.colorado.edu/wmt) is the
 web-based successor to the desktop Component Modeling Tool. WMT is a web
@@
 
-\item  \emph{Integration}. Easily hyperlink from WMT to resources on the CSDMS
+\item  \emph{Integration}. Easily hyperlink from WMT to resources on the \csdms
 portal - including model documentation, labs, lectures, tutorials and
@@
 
+\begin{figure}
+  \caption{The \csdms Web Modeling Tool}
+  \begin{center}
+    \includegraphics[scale=.25]{wmt.eps}
+  \end{center}
+  \label{fig:wmt_screenshot}
+\end{figure}
+
 \section{Digital Object Identifiers for Numerical Models}
 
-CSDMS encourages and supports
-the effort of developers to provide their numerical models as truly open source
-~\cite{ince2012case}.
-(Ince et al., 2012). All code in CSDMS is open source. Source code exposes the
-scientific hypotheses embodied in a numerical model. Solutions to a set of
-equations can take numerous forms, each with their pyramid of assumptions and
-limitations. Code transparency allows for full peer review and replication of
-results - the foundation of modern science. Code transparency also allows for
-reuse in new and clever ways, and reduces redundancy. At the same time CSDMS
-strives to ensure that model developers receive recognition for their work, even
-when code is submitted which is not (yet) described in a scientific journal.
-Therefore CSDMS has adopted the Digital Object Identifier (DOI). The DOI system
-provides a unique identification to content that is available on digital
-networks~\cite{paskin2010}. The DOI system was initiated in 1997 and contained
-mainly DOIs to publishing applications. Since 2005 DOIs were also made available
-for research data~\cite{paskin2005digital}. CSDMS appears to be the first to
-assign Digital Object Identifiers (DOIs) to numerical source code, as of 2013.
-The advantages of adopting a DOI system for models are:
-
+\csdms is a strong advocate of truly open source software.  We believe this to
+reduce redundancy, provide transparency through external review and make
+research replicable, which is fundamental to scientific practices
+~\cite{ince2012case}. \csdms acts as a venue through which open source code of
+earth surface models are promoted, either by hosting or pointing towards open
+source software. A Digital Object Identifier (DOI) system is implemented to
+guarantee recognition to open source software contributors. DOI systems are
+typically used to identify intellectual property in the digital environment.
+It is used principally by publishers, and is an implementation of the Handle
+System for persistent identifiers. The International DOI Federation (IDF)
+appoints Registration Agencies who allocate DOI prefixes, register DOI Names,
+and provide the necessary infrastructure to allow registrants to declare and
+maintain metadata. \csdms is among the first software venues to assign DOIs to
+open source software. The advantages of adopting a DOI system for open source
+software are: 
 \begin{itemize}
-
-\item Guarantee credit to a model developer.
-
-\item Easily reuse and replicate research as other researchers can directly
-      locate a referenced model.
-
-\item Providing a higher visibility, as content with a DOI is 5 times more
-      likely to deliver active links than content without.
-
-\item Giving the opportunity to for example funding agencies to track usage,
-      so to measure impact.
-
+\item Guarantee credit to the software developer
+\item Easily reuse and replicate research as software is directly locatable 
+\item Increase software visibility, DOI content is 5 times more likely to
+      deliver active links than content without.
+\item Provide e.g. funding agencies the ability to track usage, so measure
+      impact.
 \end{itemize}
 
-CSDMS collaborates with Integrated Earth Data Applications
-(IEDA), a formal Publication Agent of the DOI system through the German National
-Library of Science and Technology, to assign unique identifiers only for those
-models that contain metadata and are physically part of the CSDMS repository. An
-archive of all numerical models of the CSDMS model repository that have a DOI
-together with limited metadata as well as the source code for each model is
-provided to IEDA as a backup to guaranty access to model information and its
-source code beyond the CSDMS program; a DOI for an object is permanent, whereas
-its location and other metadata may change in future.
+\csdms requests its DOIs through the Integrated Earth Data Applications (IEDA),
+a formal Publication Agent of the DOI system that operates through the German
+National Library of Science and Technology. Only stable versions of open source
+software that are physically hosted through the \csdms repository will get a DOI
+assigned. For model software citations, \csdms recommend following the DataCite
+guidelines~\cite{brase2010datacite}. Interpreting these guidelines a citation
+should have the following construction:
 
-A new DOI is provided for each new version of a model (i.e. major upgrade of the
-source code involving a change in version number). CSDMS uses Apache Subversion,
-better known as SVN, for tracking source code versioning and revision control so
-that current and past releases and changes can be accessed through the web. As
-of March 2014, 109 models within the CSDMS model repository have a DOI. Model
-source code can be viewed as "data" and therefore CSDMS endorses citations
-defined by DataCite guidelines~\cite{brase2009datacite}. Following these guidelines, CSDMS
-strongly recommends the following structure for citing a model:
-\emph{ModelDeveloper} (\emph{PublicationYear}). \emph{ModelName}.
-\emph{ModelVersion}. \emph{Identifier}.
+\begin{shaded}
+\leftskip 0.25in
+\parindent -0.25in
+\tt{
+Developer, A., Developer, B. (Year of publication). \emph{Name of the model},
+Model Version. Identifier.
+}
+%\emph{ModelDeveloper(s)}, (\emph{PublicationYear}). \emph{ModelName},
+%\emph{ModelVersion}. \emph{Identifier}.
+\end{shaded}
+
+The DOI system for numerical models is adopted by
+\csdms as an incentive towards the earth surface processes community, to
+stimulate contributions towards open source software by ensuring recognition
+towards its developers.
 
@@
 
-The CSDMS Integration Facility operates under continuing grant 0621695 from the
+The \csdms Integration Facility operates under continuing grant 0621695 from the
 US National Science Foundation.
@@
 
+\huttonnote{Add something about model DOIs in the abstract}
+
+\huttonnote{
+\begin{itemize}
+\item improving the development process that leads to new software
+  \begin{itemize}
+  \item methods to develop sustainable software from the outset
+  \item effective approaches to reusable software created as a by-product of
+        research
+  \item impact of computer science research on the development of scientific software
+  \end{itemize}
+\end{itemize}
+}
+
+\huttonnote{This is Irina's section.}
+
+\huttonnote{
+\begin{itemize}
+\item building large and engaged user communities
+  \begin{itemize}
+  \item developing strong advocates
+  \item measurement of usage and impact
+  \end{itemize}
+\item encouraging industry's role in sustainability
+  \begin{itemize}
+  \item engagement of industry with volunteer communities
+  \item incentives for industry
+  \item incentives for community to contribute to industry-driven projects
+  \end{itemize}
+\item improving education and training
+  \begin{itemize}
+  \item best practices for providing graduate students and postdoctoral
+        researchers in domain communities with sufficient training in
+        software development
+  \item novel uses of sustainable software in education (K-20)
+  \item case studies from students on issues around software development in
+        the undergraduate or graduate curricula
+  \end{itemize}
+\end{itemize}
+}
+
+\huttonnote{This is Albert's section (he's rewriting this)}
+
+\huttonnote{
+\begin{itemize}
+\item recommending policy changes
+  \begin{itemize}
+  \item software credit, attribution, incentive, and reward
+  \item issues related to multiple organizations and multiple countries, such
+        as intellectual property, licensing, etc.
+  \item mechanisms and venues for publishing software, and the role of
+        publishers
+  \end{itemize}
+\end{itemize}
+}
+
 \end{document}

### 2014-07-14 paper.tex
Fixed brase citation.
A rough first draft.

@@
 \usepackage{fancybox}
+\usepackage{verbatim}
 
@@
 
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+
 \newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
@@
 \DeclareRobustCommand{\csdms}{\textsc{csdms}}
+\DeclareRobustCommand{\doi}{\textsc{doi}}
+\DeclareRobustCommand{\bmi}{\textsc{bmi}}
+\DeclareRobustCommand{\wmt}{\textsc{wmt}}
 
@@
   Mark D. Piper$^{\dag}$,
+  Irina Overeem$^{\dag}$
   Albert J. Kettner$^{\dag}$,
-  Irina Overeem$^{\dag}$}
+}
 
@@
 earth-system models and tools to an international (67 country) community
-with the goal of building a complete set of tools necessary to model the
+with the goal of building the set of tools necessary to model the
 earth system. \csdms is a place modelers look to for access to hundreds open
@@
 testing, continious integration, test-driven development, well-written
-\emph{clean code}.
+\emph{clean code} are all topics of the educational mission of \csdms{}.
 
@@
 
-The Community Surface Dynamics Modeling System (\csdms{})
-~\cite{peckham2012component}
-is an NSF-funded project that began in 2007. Its mission is to help a diverse
-community of surface dynamics model \emph{developers} and model \emph{users}
-work together toward
-common goals and standards. Part of this effort involves creating a repository
-of open-source models. Another part involves converting models to plug-and-play
-components that can be reused via dynamic linking within the \csdms framework to
-create new models. In building its modeling framework, \csdms has leveraged
+The mission of the Community Surface Dynamics Modeling System (\csdms{})
+~\cite{peckham2012component} is to help a diverse modeling community toward
+common goals and standards.
+%The Community Surface Dynamics Modeling System (\csdms{})
+%is an NSF-funded project that began in 2007. Its mission is to help a diverse
+%community of surface dynamics model \emph{developers} and model \emph{users}
+%work together toward
+%common goals and standards.
+This effort involves creating:
+\begin{itemize}
+  \item a repository of source code and metadata for open-source models and
+    tools (section~\ref{sec:organizing})
+  \item reusable plug-and-play model components and a framework
+        within which these components can be coupled to create new models
+        (section~\ref{sec:reusable})
+  \item an efficient and open modeling community through education and
+        standards (section~\ref{sec:education})
+\end{itemize}
+
+%of open-source models. Another part involves converting models to plug-and-play
+%components that can be reused via dynamic linking within the \csdms framework to
+%create new models.
+
+In building a modeling framework, \csdms{} has leveraged
 several existing, well-established and open-source software tools. For example,
-\csdms uses three tools from the Common Component Architecture (CCA) toolchain:
-Babel, Bocca and Ccaffeine. Babel provides interoperability between components
+\csdms{} uses tools from the Common Component Architecture (CCA)
+~\cite{armstrong1999toward} toolchain: \emph{Babel}, and \emph{Bocca}. Babel
+provides interoperability between components
 written in different languages; it currently supports C, C++, Fortran, Java,
 and Python. Bocca helps with creating CCA-compliant components and managing CCA
-component projects. Ccaffeine is a CCA-compliant framework for linking
-components into new applications. All three tools were designed to enable
-component-based modeling in an HPC environment, but they can also be used on
-desktop systems. \csdms has developed innovative model/component interfaces for
-use with the CCA tools, including the Basic Model Interface (BMI), which uses
-the \csdms Standard Names (both described below) and the framework-level
-Component Model Interface (CMI). \csdms has also developed a web-based client
-for its modeling framework (the \csdms Web Modeling Tool, WMT, described below)
-that allows users to compose new models by connecting and configuring
-components through a simple, browser-based graphical interface.
+component projects.
+%Ccaffeine is a CCA-compliant framework for linking
+%components into new applications. All three tools were designed to enable
+%component-based modeling in an HPC environment, but they can also be used on
+%desktop systems.
 
+\csdms{} has developed innovative model/component interfaces that promote
+model reuse and interoperability,
+including the Basic Model Interface (\bmi{} - section~\ref{sec:bmi}), which
+uses
+the \csdms{} Standard Names (section~\ref{sec:standardnames}) and a framework
+within which models can be coupled and enhanced with a set of \bmi{}-compatible
+service components.
+%Component Model Interface (CMI).
+The \csdms{} Web Modeling Tool (\wmt{} - section~\ref{sec:wmt}) is a web-based client that provides a
+graphical interface to this framework and allows users to compose new models
+by connecting and configuring components in a simple, browser-based graphical
+interface.
+
+%\csdms{} has also developed a web-based client through which users 
+%for its modeling framework (the \csdms Web Modeling Tool, WMT, described below)
+%that allows users to compose new models by connecting and configuring
+%components through a simple, browser-based graphical interface.
+
+\begin{comment}
 In order to realize the full power and benefits of component-based modeling, a
-modeling framework like \csdms needs an efficient mechanism to convert as many
+modeling framework like \csdms{} needs an efficient mechanism to convert as many
 open-source models as possible to reusable plug-and-play components. Since this
-necessarily requires some involvement from each model's developer, this
-mechanism must be designed to:
-
+necessarily requires some involvement from each model's developer (\emph{who,
+likely, is not paid to do this}), this mechanism must require minimal effort
+and the benifit to the developer should be clear.
+In the first few years of the \csdms{} project, we worked directly with model
+developers and conducted one-on-one interviews to hear their concerns and
+complaints about early designs and other frameworks. In addition to being easy
+to implement, the following requirements of a model wrapping procedure became
+clear:
+%Through working directly with model developers and listening to their concerns
+%and complaints about early designs and other frameworks,
+%the following requirements presented themselves:
 \begin{itemize}
 
-\item require minimal effort from the model developer,
+%\item require minimal effort from the model developer,
 
@@
 
-These requirements became clear during the first few years of the \csdms project
-by working directly with model developers and listening to their concerns and
-complaints about early designs and other frameworks.
+%These requirements became clear during the first few years of the \csdms{}
+%project
+%by working directly with model developers and listening to their concerns and
+%complaints about early designs and other frameworks.
+\end{comment}
 
-\section{Building a User Community}
+\section{Organizing and Documenting Open-Source Models}
+\label{sec:organizing}
 
-\csdms{} provides cyber-infrastructure to promote the quantitative modeling of
-earth surface processes. The US National Science Foundation funds CSDMS, but
-the community is international and includes industry and federal agency
-representatives in addition to members from academic institutions. The
-initiative has been growing with >150 people per year. CSDMS now fosters a
-community of ~1150 scientists who work on prediction of the movement of fluids,
-and the production, erosion, transport, and deposition of sediment and
-nutrients in landscapes and seascapes.
+\subsection{The \csdms{} Model Repository}
+\label{sec:repository}
 
-\section{Model Repository}
+Open-source software reduces redundancy, provides transparency through
+external review and makes research replicable, which is fundamental to
+scientific practices~\cite{ince2012case}.
+Software accessibility is a key value of the \csdms{} mission. \csdms{}
+ensures code developed by individual researchers or small research teams
+remains accessible beyond the lifetime of a specific research project. Code
+developers are required to distribute their codes under open-source licenses.
 
-Software accessibility is a key value of the CSDMS mission.
-CSDMS ensures that code developed by individual researchers or small research teams remains accessible beyond the lifetime of a specific research project.
-Code developers are coaxed into providing their codes under open-source licenses.
+\csdms{} has built an online model repository that not only contains model
+source code but also documents the submitted code as model metadata, runs
+basic smoke-tests of the models, (optionally) maintains the source code in a
+\csdms-hosted version control system, and provides easy downloads of the
+original source. The repository now contains over 200 models and tools (over
+4M lines of code) and has recieved over 13k model downloads.
 
-CSDMS has built an online model repository that importantly documents the submitted code with metadata, smoke-tests codes on the CSDMS High Performance Computing Cluster, maintains a snapshot of submitted source code in a version-control system, and provides easy downloads of the original source code.
+%importantly documents the
+%submitted code with metadata, smoke-tests codes on the CSDMS High Performance Computing Cluster, maintains a snapshot of submitted source code in a version-control system, and provides easy downloads of the original source code.
 This repository now contains 172 codes and 56 tools, or 4,074,236 lines of source code (a source code line is defined as any line that is not part of a comment, and is not blank).
 
-We do know our community values this resource, we track downloads of each individual code and the total number of downloads since November 2008 is >13,218 (as of July 2014).
+%We do know our community values this resource, we track downloads of each individual code and the total number of downloads since November 2008 is >13,218 (as of July 2014).
 
-\section{Software sustainability through expanded functionality}
+\subsection{Digital Object Identifiers for Numerical Models}
+\label{sec:doi}
 
-Another important goal of CSDMS is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales.
-Interdisciplinarity and coupling is a critical step towards advancing predictive modeling~\cite{voinov2010community}.
+%\csdms{} is a strong advocate of truly open source software.
 
-CSDMS develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.
+%The \doi{} system for numerical models is adopted by
+%\csdms as an incentive towards the earth surface processes community, to
+%stimulate contributions towards open source software by ensuring recognition
+%towards its developers.
 
-The modeling coupling framework is designed to catalyze Earth-surface process research by empowering a broad community of scientists, students and applied modelers with computing tools, streamlining the process of idea generation and hypothesis testing through linked surface dynamics models, and enabling rapid creation and coupling of models tailored to specific settings, scientific problems, and time scales
+\csdms{} has implemented a
+Digital Object Identifier (\doi{}) system for software within the model
+repository to guarantee recognition to software contributors.
+\csdms{} was among the first software venues to assign \doi{}s to
+open source software. The advantages of adopting a \doi{} system for open
+source software are: 
+\begin{itemize}
+\item Guarantee credit to the software developer
+\item Easily reuse and replicate research as software is directly locatable 
+\item Increase software visibility, \doi{} content is 5 times more likely to
+      deliver active links than content without.
+\item Provide e.g. funding agencies the ability to track usage, so measure
+      impact.
+\end{itemize}
 
-\section{CSDMS has an educational mission}
+%\csdms{} acts as a venue through which open source code of
+%earth surface models are promoted, either by hosting or pointing towards open
+%source software.
 
-WMT provides the ability to run models with an easy to use graphical interface and intends to serve novice and more experienced modelers alike.
-The WMT beta version was released in May 2014  as a sequel to the CMT.
-IF staff habitually evaluates the effectiveness of this tool and uses the feedback to improve the functionality: modeling courses at the University of Colorado, clinics at the International Summer Institute at the University of Minnesota, and clinics at annual CSDMS Annual Meetings are the dominant forums for this testing.
+%A Digital Object Identifier (\doi{}) system is implemented to
+%guarantee recognition to open source software contributors. \doi{} systems are
+%typically used to identify intellectual property in the digital environment.
+%It is used principally by publishers, and is an implementation of the Handle
+%System for persistent identifiers. The International \doi{} Federation (IDF)
+%appoints Registration Agencies who allocate \doi{} prefixes, register \doi{} Names,
+%and provide the necessary infrastructure to allow registrants to declare and
+%maintain metadata.
+%\csdms is among the first software venues to assign \doi{}s to
+%open source software. The advantages of adopting a \doi{} system for open source
+%software are: 
+%\begin{itemize}
+%\item Guarantee credit to the software developer
+%\item Easily reuse and replicate research as software is directly locatable 
+%\item Increase software visibility, \doi{} content is 5 times more likely to
+%      deliver active links than content without.
+%\item Provide e.g. funding agencies the ability to track usage, so measure
+%      impact.
+%\end{itemize}
 
-Over the last 4 years >150 graduate students were enrolled in intensive modeling courses, originating from geography, hydrology, geology, oceanography, earth sciences and engineering departments.
-These courses provided an excellent use case for us to test usability and access to the tool.
-Students reported in post-course surveys that they were unfamiliar with the use of models on a high performance computing cluster beforehand (rated their familiarity as 1.5 average on a scale of 1-5), but ended the courses with familiarity and being more comfortable with running and changing simulations as well as visualizing results (average rating approaches 4 on a Likert scale of 1-5).
+\begin{comment}
+\csdms{} requests its \doi{}s through the Integrated Earth Data Applications,
+%(\textsc{IEDA}),
+a formal Publication Agent of the \doi{} system that operates through the German
+National Library of Science and Technology.
+\end{comment}
 
-Focused hands-on clinics at meetings also allowed more than 150 CSDMS members to experience the ease of use of CMT, and now WMT, and try several examples of how to use the tool.
+Only stable versions of open source
+software that are physically hosted through the \csdms{} repository will be
+assigned a \doi{}. For model software citations, \csdms{} recommends the
+following the DataCite guidelines~\cite{brase2009datacite}.
 
-Course and clinic materials are posted online as educational resources.
-IF staff are developing WMT-specific help pages for all models currently available as components, so that users can easily learn about model parameters and input files.
-In addition, IF staff members are developing web-based video tutorials that show how to run models.
+\begin{shaded}
+\leftskip 0.25in
+\parindent -0.25in
+\tt{
+Developer, A., Developer, B. (Year of publication). \emph{Name of the model},
+Model Version. Identifier.
+}
+%\emph{ModelDeveloper(s)}, (\emph{PublicationYear}). \emph{ModelName},
+%\emph{ModelVersion}. \emph{Identifier}.
+\end{shaded}
 
-\section{The Basic Modeling Interface}
+\section{Reusable Components}
+\label{sec:reusable}
 
-The \emph{Basic Modeling Interface} (BMI) is a component-coupling library
+\subsection{Sustainability through standards}
+
+\subsubsection{The Basic Modeling Interface}
+\label{sec:bmi}
+
+The \emph{Basic Modeling Interface}\footnote{https://github.com/csdms/bmi}
+(\bmi{}) is a component-coupling library
 interface specification designed by \csdms~\cite{peckham2012component,
 syvitski2014plug}.  In this context, a component is software that models a
-particular environmental process and can be "plugged into" another component.
+particular environmental process and can be \emph{plugged into} another
+component.
 
-The BMI specification was designed without any particular model-coupling
+The \bmi{} specification was designed without any particular model-coupling
 framework in mind.  Rather, it was designed to be framework agnostic. We
@@
 
-The BMI identifies the entry points into software components that provides a
+The \bmi{} identifies the entry points into software components that provides a
 calling appication with the necessary level of control to connect components.
@@
 provides functionality to \emph{initialize}, \emph{update}, and
-\emph{finalize} a component. However, for more tight coupling the components
-must implement an interface to provide data access, and model metadata.
+\emph{finalize} a component. However, for more complicated (and 2-way)
+coupling components must implement the \bmi{} methods providing data access,
+and model metadata.
 
-\csdms has found that BMI is an acceptable target for model developers. The use
-of BMI has also dramatically reduced the effort required by \csdms staff to
-create and maintain components. So long as a model component strictly exposes
-a BMI, it can automatically be incorporated into the \csdms model-coupling
-framework where it can then connect to compatible components to form larger
-models.
+\csdms{} has found that \bmi{} is an acceptable target for model developers.
+The use of \bmi{} has also dramatically reduced the effort required by
+\csdms{} staff to create and maintain components. So long as a model component
+strictly exposes a \bmi{}, it can automatically be incorporated into the
+\csdms{} model-coupling framework where it can then connect to compatible
+components to form larger models and obtain additional functionality through
+\bmi{} compatible tools (for example, NetCDF output, grid mapping).
 
@@
 
-\section{CSDMS Standard Names}
+\subsubsection{CSDMS Standard Names}
+\label{sec:standardnames}
 
-\huttonnote{Taken from the SSI proposal. I need to rewrite this.}
+%\huttonnote{Taken from the SSI proposal. I need to rewrite this.}
 
-The \csdms Standard Names are a lingua franca for variable names between model
-components. They play an important role in the BMI as they provide a mapping of
-a model's internal variable names to a common language used by the BMI getter
-and setter functions.
+The \csdms{} \emph{Standard Names}\footnote{https://github.com/csdms/bmi}
+are a common language for variable names exchanged between models. They play
+an important role in the \bmi{} as they provide a mapping of a model's
+internal variable names to a common language used by the \bmi{} getter and
+setter functions.
 
-Most models require input variables and produce output variables. In a
-component-based modeling framework, like \csdms, a set of components becomes a
-complete model when every component is able to obtain the input variables it
-needs from another component. The \csdms Standard Names were developed to
-provide a practical solution to this semantic mediation problem
-~\cite{peckham2012component, syvitski2014plug}.
-While the CF Convention Standard Names, which were
-introduced in the domain of the ocean and atmosphere model domains, have
-somewhat overlapping goals, the \csdms Standard Names provide a more
-comprehensive set of naming rules and patterns for creating unique labels for
-model variables that are not specific to any particular modeling domain.
+Most models require input variables and produce output variables. For model
+components to be reusable and interoperable (either within the \csdms{}
+framework or any framework in general)
+%In a component-based modeling framework, like \csdms{},
+a set of components become a complete model when every component is able to
+obtain the input variables it needs from another component. The \csdms{}
+Standard Names were developed to provide a practical solution to this
+semantic mediation problem ~\cite{peckham2012component, syvitski2014plug}.
+%While the CF Convention Standard Names, which were
+%introduced in the domain of the ocean and atmosphere model domains, have
+%somewhat overlapping goals,
+The \csdms{} Standard Names provide a comprehensive set of naming rules and
+patterns for creating unique labels for model variables that are not specific
+to any particular modeling domain.
 
@@
 
-\section{Reusable Components}
+\subsection{Connecting and running models through the web}
+\label{sec:wmt}
 
-\subsection{The \csdms Modeling Framework}
+%\huttonnote{Taken from the SSI proposal. I need to rework this a bit.}
 
-\subsection{Connecting Components with the Web Modeling Tool}
-
-\huttonnote{Taken from the SSI proposal. I need to rework this a bit.}
-
-The \csdms Web Modeling Tool (WMT; https://csdms.colorado.edu/wmt) is the
-web-based successor to the desktop Component Modeling Tool. WMT is a web
-application that provides an Ajax client-side graphical interface (the WMT
-client) and a RESTful server-side database and API (the WMT server) that allows
-users to build and run coupled Earth system models on a high-performance
-computing cluster (HPCC) from a web browser.
-WMT was designed with four objectives:
+The \csdms{} Web Modeling Tool\footnote{https://csdms.colorado.edu/wmt}
+(\wmt{}) is a web
+%is the web-based successor to the desktop Component Modeling Tool. \wmt{} 
+application that provides an Ajax client-side (the \wmt{}
+client\footnote{https://github.com/csdms/wmt-client}) graphical interface
+(figure~\ref{fig:wmt_screenshot}) and a RESTful
+server-side database and API (the \wmt{} server\footnote{https://github.com/csdms/wmt}) that allows users to build and run coupled Earth system models on
+a high-performance computing cluster from a web browser. \wmt{} was designed
+with four objectives:
 \begin{itemize}
@@
 \item \emph{Accessibility}. As a web-based application, if you have access
-to the Internet, you have access to WMT.
+      to the Internet, you have access to \wmt{}.
 
-\item  \emph{Integration}. Easily hyperlink from WMT to resources on the \csdms
-portal - including model documentation, labs, lectures, tutorials and
-movies - or to other resources on the Internet.
+\item \emph{Integration}. Easily hyperlink from \wmt{} to resources on the 
+      \csdms{} modeling portal (model documentation, labs, lectures,
+      tutorials)
+      %tutorials and movies - or to other resources on the Internet.
 
-\item \emph{Portability}. WMT has a native JavaScript interface, so it can be
-accessed on any modern web browser, including tablet and mobile
+\item \emph{Portability}. \wmt{} has a native JavaScript interface, so it can
+      be accessed on any modern web browser
+%, including tablet and mobile
       versions of browsers.
 
-\item \emph{Maintenance}. Because modern browsers tend to adhere to web
-standards, which lead to fewer cross-compatibility issues than
-operating systems, only one version of WMT needs to be developed
-and maintained.
+%\item \emph{Maintenance}. Because modern browsers tend to adhere to web
+%      standards, which lead to fewer cross-compatibility issues than
+%      operating systems, only one version of WMT needs to be developed
+%      and maintained.
 
 \end{itemize}
+With \wmt{}, users can: \emph{run} single component models, build (2-way)
+\emph{coupled models}, view and \emph{edit} model parameters, \emph{share}
+saved models with others in the community.
 
-With WMT, a user can:
-
+\begin{comment}
 \begin{itemize}
 
-\item select a Common Component Architecture (CCA) component model
-      from a list to run in standalone mode;
+\item select and run a single component model
 
-\item build a coupled model from multiple CCA components organized as nodes
-      of a tree structure;
+\item build a coupled model from multiple components
 
@@
 
-\item save models to a server, where they can be accessed on any computer
-      connected to the Internet;
+%\item save models to a server, where they can be accessed on any computer
+%      connected to the Internet;
 
@@
 
-\item run a model by connecting to a remote HPCC where the components are
-      installed.
+%\item run a model by connecting to a remote HPCC where the components are
+%      installed.
 
 \end{itemize}
+\end{comment}
 
-Although WMT is web-based, the building and configuring of a model can be done
-offline. Reconnection is necessary only when saving a model and submitting it
-for a run.
+%Although WMT is web-based, the building and configuring of a model can be done
+%offline. Reconnection is necessary only when saving a model and submitting it
+%for a run.
+%\section{Reusable Components}
+\begin{comment}
+\subsection{Software sustainability through expanded functionality}
+
+Another important goal of CSDMS is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales.
+Interdisciplinarity and coupling is a critical step towards advancing predictive modeling~\cite{voinov2010community}.
+
+CSDMS develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.
+
+The modeling coupling framework is designed to catalyze Earth-surface process research by empowering a broad community of scientists, students and applied modelers with computing tools, streamlining the process of idea generation and hypothesis testing through linked surface dynamics models, and enabling rapid creation and coupling of models tailored to specific settings, scientific problems, and time scales
+\end{comment}
+
+%\section{The \csdms Modeling Framework}
+
+\section{Education and Evaluation of \csdms{} tools}
+\label{sec:education}
+
+%\subsection{CSDMS has an educational mission}
+%\subsection{Education and Evaluation}
+
+%\wmt{} provides the ability to run models with an easy to use graphical
+%interface and intends to serve novice and more experienced modelers alike.
+%The WMT beta version was released in May 2014  as a sequel to the CMT.
+
+\csdms{} constantly evaluates the effectiveness of its tools and standards
+through user feedback from clinics and tutorials and one-on-one interviews
+with \csdms{} members.  Modeling courses at the University of Colorado,
+clinics at the International Summer Institute at the University of Minnesota,
+and clinics at annual CSDMS Annual Meetings are the dominant forums for
+this testing and feedback.
+
+%IF staff habitually evaluates the effectiveness of this tool and uses the feedback to improve the functionality: modeling courses at the University of Colorado, clinics at the International Summer Institute at the University of Minnesota, and clinics at annual CSDMS Annual Meetings are the dominant forums for this testing.
+
+Over the last 4 years over 150 graduate students - from geography, hydrology,
+geology, oceanography, earth sciences and engineering departments - have
+enrolled in intensive \csdms{}-run modeling courses. These courses
+provide an excellent use case for us to test usability and access to the
+tool. Students reported in post-course surveys that they were unfamiliar with
+the use of models on a high performance computing cluster beforehand (rated
+their familiarity as 1.5 average on a scale of 1-5), but ended the courses
+with familiarity and being more comfortable with running and changing
+simulations as well as visualizing results (average rating approaches 4 on
+a Likert scale of 1-5).
+
+%Focused hands-on clinics at meetings also allowed more than 150 CSDMS members to experience the ease of use of \wmt{}, and try several examples of how to use the tool.
+
+Course and clinic materials are posted online as educational resources.
+\csdms{} develops \wmt{}-specific help pages for all models currently
+available as components, so that users can easily learn about model
+parameters and input files. In addition, we are developing web-based video
+tutorials that show how to run models.
+
+\section{Summary: Building an active and efficient Modeling Community}
+
+\csdms{} provides cyber-infrastructure to promote the quantitative modeling of
+earth surface processes. The US National Science Foundation funds \csdms{}, but
+the community is international and includes industry and federal agency
+representatives in addition to members from academic institutions. The
+initiative has been growing with more than 150 people per year. \csdms{} now
+fosters a community of approximately 1200 scientists who work on prediction
+of the movement of fluids, and the production, erosion, transport, and
+deposition of sediment and nutrients in landscapes and seascapes.
+
+\section{Acknowledgements}
+
+The \csdms{} Integration Facility operates under continuing grant 0621695
+from the US National Science Foundation.
 
@@
 
-\section{Digital Object Identifiers for Numerical Models}
-
-\csdms is a strong advocate of truly open source software.  We believe this to
-reduce redundancy, provide transparency through external review and make
-research replicable, which is fundamental to scientific practices
-~\cite{ince2012case}. \csdms acts as a venue through which open source code of
-earth surface models are promoted, either by hosting or pointing towards open
-source software. A Digital Object Identifier (DOI) system is implemented to
-guarantee recognition to open source software contributors. DOI systems are
-typically used to identify intellectual property in the digital environment.
-It is used principally by publishers, and is an implementation of the Handle
-System for persistent identifiers. The International DOI Federation (IDF)
-appoints Registration Agencies who allocate DOI prefixes, register DOI Names,
-and provide the necessary infrastructure to allow registrants to declare and
-maintain metadata. \csdms is among the first software venues to assign DOIs to
-open source software. The advantages of adopting a DOI system for open source
-software are: 
-\begin{itemize}
-\item Guarantee credit to the software developer
-\item Easily reuse and replicate research as software is directly locatable 
-\item Increase software visibility, DOI content is 5 times more likely to
-      deliver active links than content without.
-\item Provide e.g. funding agencies the ability to track usage, so measure
-      impact.
-\end{itemize}
-
-\csdms requests its DOIs through the Integrated Earth Data Applications (IEDA),
-a formal Publication Agent of the DOI system that operates through the German
-National Library of Science and Technology. Only stable versions of open source
-software that are physically hosted through the \csdms repository will get a DOI
-assigned. For model software citations, \csdms recommend following the DataCite
-guidelines~\cite{brase2010datacite}. Interpreting these guidelines a citation
-should have the following construction:
-
-\begin{shaded}
-\leftskip 0.25in
-\parindent -0.25in
-\tt{
-Developer, A., Developer, B. (Year of publication). \emph{Name of the model},
-Model Version. Identifier.
-}
-%\emph{ModelDeveloper(s)}, (\emph{PublicationYear}). \emph{ModelName},
-%\emph{ModelVersion}. \emph{Identifier}.
-\end{shaded}
-
-The DOI system for numerical models is adopted by
-\csdms as an incentive towards the earth surface processes community, to
-stimulate contributions towards open source software by ensuring recognition
-towards its developers.
-
-\section{Acknowledgements}
-
-The \csdms Integration Facility operates under continuing grant 0621695 from the
-US National Science Foundation.
-
 \bibliographystyle{plain}
@@
 
+\begin{comment}
 \huttonnote{Add something about model DOIs in the abstract}
@@
 }
+\end{comment}
 

### 2014-07-14 paper.tex
Minor changes.

@@
 
-\title[]{Building Sustainable Software - The CSDMS-2.0 Approach}
+\title[]{Building Sustainable Software - The CSDMS Approach}
 
@@
 
-The Community Surface Dynamics Modeling System (\csdms{}) is an NSF funded
+\csdms, The Community Surface Dynamics Modeling System, is an NSF funded
 project whose focus is to aid a diverse community of earth and ocean system
@@
 with the goal of building the set of tools necessary to model the
-earth system. \csdms is a place modelers look to for access to hundreds open
+earth system. \csdms{} is a place modelers look to for access to hundreds open
 source surface-dynamics models and tools, as well as model metadata. Such a
@@
 duplication of effors by presenting the current state of modeling efforts.
-To increase software sustainability, composability and interoperability \csdms
+To increase software sustainability, composability and interoperability \csdms{}
 promotes standards that define common modeling interfaces, semantic mediation
 between models, and model metadata. Through online resources and workshops,
-\csdms promotes software engineering best practices, which are foreign to many
+\csdms{} promotes software engineering best practices, which are foreign to many
 developers within our modeling community. For example, version control, unit
@@
   \item a repository of source code and metadata for open-source models and
-    tools (section~\ref{sec:organizing})
+    tools (Section~\ref{sec:organizing})
   \item reusable plug-and-play model components and a framework
         within which these components can be coupled to create new models
-        (section~\ref{sec:reusable})
+        (Section~\ref{sec:reusable})
   \item an efficient and open modeling community through education and
-        standards (section~\ref{sec:education})
+        standards (Section~\ref{sec:education})
 \end{itemize}
@@
 model reuse and interoperability,
-including the Basic Model Interface (\bmi{} - section~\ref{sec:bmi}), which
+including the Basic Model Interface (\bmi{} - Section~\ref{sec:bmi}), which
 uses
-the \csdms{} Standard Names (section~\ref{sec:standardnames}) and a framework
+the \csdms{} Standard Names (Section~\ref{sec:standardnames}) and a framework
 within which models can be coupled and enhanced with a set of \bmi{}-compatible
@@
 %Component Model Interface (CMI).
-The \csdms{} Web Modeling Tool (\wmt{} - section~\ref{sec:wmt}) is a web-based client that provides a
+The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based client that provides a
 graphical interface to this framework and allows users to compose new models
@@
 
-\subsection{Sustainability through standards}
+\subsection{Sustainability and Interoperability Through Standards}
 
@@
 
-\subsubsection{CSDMS Standard Names}
+\subsubsection{\csdms{} Standard Names}
 \label{sec:standardnames}
@@
 
-\subsection{Connecting and running models through the web}
+\subsection{Connecting and Running Models Through the Web}
 \label{sec:wmt}
@@
 
-\section{Education and Evaluation of \csdms{} tools}
+\section{Education and Evaluation of \csdms{} Tools}
 \label{sec:education}
@@
 
-\section{Summary: Building an active and efficient Modeling Community}
+\section{Summary: Building an active and efficient modeling community}
 

### 2014-07-14 paper.tex
Cited overeem et al.

@@
         (Section~\ref{sec:reusable})
-  \item an efficient and open modeling community through education and
-        standards (Section~\ref{sec:education})
+  \item an efficient and open modeling community through standards (Section
+        ~\ref{sec:standards}) and education (Section~\ref{sec:education})
 \end{itemize}
@@
 \subsection{Sustainability and Interoperability Through Standards}
+\label{sec:standards}
 
@@
 
-The \csdms{} \emph{Standard Names}\footnote{https://github.com/csdms/bmi}
+The \csdms{} \emph{Standard Names}
+\footnote{https://github.com/csdms/standard\_names}
 are a common language for variable names exchanged between models. They play
@@
 simulations as well as visualizing results (average rating approaches 4 on
-a Likert scale of 1-5).
+a Likert scale of 1-5)~\cite{overeem2013strategies}.
 

### 2014-07-14 paper.tex
Cleaned up excess comments.

@@
 
-%\usepackage{multibbl}
-%\usepackage{bibtopic}
-
 \usepackage{color}
@@
 \usepackage{graphicx}
-%\usepackage{graphics}
 \usepackage[latin9]{inputenc}
@@
 \usepackage{verbatim}
-
-%\setcounter{secnumdepth}{3}
-%\setcounter{tocdepth}{3}
 
@@
 \newcommand{\huttonnote}[1]{ {\textcolor{magenta}    { ***Eric:      #1 }}}
+
 \DeclareRobustCommand{\csdms}{\textsc{csdms}}
@@
 ~\cite{peckham2012component} is to help a diverse modeling community toward
-common goals and standards.
-%The Community Surface Dynamics Modeling System (\csdms{})
-%is an NSF-funded project that began in 2007. Its mission is to help a diverse
-%community of surface dynamics model \emph{developers} and model \emph{users}
-%work together toward
-%common goals and standards.
-This effort involves creating:
+common goals and standards. This effort involves creating:
 \begin{itemize}
@@
   \item reusable plug-and-play model components and a framework
-        within which these components can be coupled to create new models
+        within which they can be coupled to create new models
         (Section~\ref{sec:reusable})
-  \item an efficient and open modeling community through standards (Section
-        ~\ref{sec:standards}) and education (Section~\ref{sec:education})
+  \item an \emph{efficient} and \emph{open} modeling community through
+        standards (Section ~\ref{sec:standards}) and education
+        (Section~\ref{sec:education})
 \end{itemize}
-
-%of open-source models. Another part involves converting models to plug-and-play
-%components that can be reused via dynamic linking within the \csdms framework to
-%create new models.
 
@@
 component projects.
-%Ccaffeine is a CCA-compliant framework for linking
-%components into new applications. All three tools were designed to enable
-%component-based modeling in an HPC environment, but they can also be used on
-%desktop systems.
 
@@
 service components.
-%Component Model Interface (CMI).
 The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based client that provides a
@@
 interface.
-
-%\csdms{} has also developed a web-based client through which users 
-%for its modeling framework (the \csdms Web Modeling Tool, WMT, described below)
-%that allows users to compose new models by connecting and configuring
-%components through a simple, browser-based graphical interface.
 
@@
 
-%importantly documents the
-%submitted code with metadata, smoke-tests codes on the CSDMS High Performance Computing Cluster, maintains a snapshot of submitted source code in a version-control system, and provides easy downloads of the original source code.
-This repository now contains 172 codes and 56 tools, or 4,074,236 lines of source code (a source code line is defined as any line that is not part of a comment, and is not blank).
-
-%We do know our community values this resource, we track downloads of each individual code and the total number of downloads since November 2008 is >13,218 (as of July 2014).
-
 \subsection{Digital Object Identifiers for Numerical Models}
 \label{sec:doi}
-
-%\csdms{} is a strong advocate of truly open source software.
-
-%The \doi{} system for numerical models is adopted by
-%\csdms as an incentive towards the earth surface processes community, to
-%stimulate contributions towards open source software by ensuring recognition
-%towards its developers.
 
@@
 
-%\csdms{} acts as a venue through which open source code of
-%earth surface models are promoted, either by hosting or pointing towards open
-%source software.
-
-%A Digital Object Identifier (\doi{}) system is implemented to
-%guarantee recognition to open source software contributors. \doi{} systems are
-%typically used to identify intellectual property in the digital environment.
-%It is used principally by publishers, and is an implementation of the Handle
-%System for persistent identifiers. The International \doi{} Federation (IDF)
-%appoints Registration Agencies who allocate \doi{} prefixes, register \doi{} Names,
-%and provide the necessary infrastructure to allow registrants to declare and
-%maintain metadata.
-%\csdms is among the first software venues to assign \doi{}s to
-%open source software. The advantages of adopting a \doi{} system for open source
-%software are: 
-%\begin{itemize}
-%\item Guarantee credit to the software developer
-%\item Easily reuse and replicate research as software is directly locatable 
-%\item Increase software visibility, \doi{} content is 5 times more likely to
-%      deliver active links than content without.
-%\item Provide e.g. funding agencies the ability to track usage, so measure
-%      impact.
-%\end{itemize}
-
-\begin{comment}
-\csdms{} requests its \doi{}s through the Integrated Earth Data Applications,
-%(\textsc{IEDA}),
-a formal Publication Agent of the \doi{} system that operates through the German
-National Library of Science and Technology.
-\end{comment}
-
 Only stable versions of open source
@@
 }
-%\emph{ModelDeveloper(s)}, (\emph{PublicationYear}). \emph{ModelName},
-%\emph{ModelVersion}. \emph{Identifier}.
 \end{shaded}
@@
 
-%The introduction of the
-%\csdms Standard Names provides a practical and elegant solution to the semantic
-%mediation problem and Babel solves the language interoperability problem.
-%\csdms has also assembled a large repository of surface process models that now
-%includes over 171 open-source models and 56 tools. The repository contains
-%terrestrial, coastal, marine, hydrological, carbonate and atmospheric models.
-%Geodynamic, atmospheric and ecosystem models will be added over the next few
-%years.
-
-%So far, 55 of the 171 models have been wrapped for use as reusable
-%plug-and-play components (although only 6 are currently visible in the new WMT).
-
 \subsubsection{\csdms{} Standard Names}
 \label{sec:standardnames}
-
-%\huttonnote{Taken from the SSI proposal. I need to rewrite this.}
 
@@
 framework or any framework in general)
-%In a component-based modeling framework, like \csdms{},
 a set of components become a complete model when every component is able to
@@
 semantic mediation problem ~\cite{peckham2012component, syvitski2014plug}.
-%While the CF Convention Standard Names, which were
-%introduced in the domain of the ocean and atmosphere model domains, have
-%somewhat overlapping goals,
 The \csdms{} Standard Names provide a comprehensive set of naming rules and
@@
 
-%The \csdms Standard Names are currently being extended to solid Earth modeling
-%domains, including geodynamics, seismology, electromagnetic induction and
-%petrology, in the framework of the "Earth System Bridge" EarthCube project led
-%by PI Peckham. They also serve as the base for the metadata standard for
-%numerical models (MMF) which is currently in development within \csdms.
-
-%We are acutely aware of multiple other efforts to provide cross-domain
-%ontologies in the realm of geosciences. PIs Peckham and Kelbert are involved
-%with the initial steps in semantic crosswalk development recently initialized
-%within the EarthCube community that aim to achieve synergy between these
-%currently divergent metadata ontologies.
-
 \subsection{Connecting and Running Models Through the Web}
@@
 
-%\huttonnote{Taken from the SSI proposal. I need to rework this a bit.}
-
 The \csdms{} Web Modeling Tool\footnote{https://csdms.colorado.edu/wmt}
 (\wmt{}) is a web
-%is the web-based successor to the desktop Component Modeling Tool. \wmt{} 
 application that provides an Ajax client-side (the \wmt{}
@@
       be accessed on any modern web browser
-%, including tablet and mobile
-      versions of browsers.
 
@@
 
-%Although WMT is web-based, the building and configuring of a model can be done
-%offline. Reconnection is necessary only when saving a model and submitting it
-%for a run.
-%\section{Reusable Components}
 \begin{comment}
@@
 
-%\section{The \csdms Modeling Framework}
-
 \section{Education and Evaluation of \csdms{} Tools}
 \label{sec:education}
-
-%\subsection{CSDMS has an educational mission}
-%\subsection{Education and Evaluation}
 
@@
 
-%IF staff habitually evaluates the effectiveness of this tool and uses the feedback to improve the functionality: modeling courses at the University of Colorado, clinics at the International Summer Institute at the University of Minnesota, and clinics at annual CSDMS Annual Meetings are the dominant forums for this testing.
-
 Over the last 4 years over 150 graduate students - from geography, hydrology,
@@
 a Likert scale of 1-5)~\cite{overeem2013strategies}.
-
-%Focused hands-on clinics at meetings also allowed more than 150 CSDMS members to experience the ease of use of \wmt{}, and try several examples of how to use the tool.
 

### 2014-07-14 paper.tex
Added CU to affiliation.
Moved WMT screenshot figure into WMT section.

@@
 
-\thanks{{}$^{\dag}$ Community Surface Dynamics Modeling System, Boulder, CO}
+\thanks{{}$^{\dag}$Community Surface Dynamics Modeling System, University of
+Colorado}
 
@@
 client\footnote{https://github.com/csdms/wmt-client}) graphical interface
-(figure~\ref{fig:wmt_screenshot}) and a RESTful
+(Figure~\ref{fig:wmt_screenshot}) and a RESTful
 server-side database and API (the \wmt{} server\footnote{https://github.com/csdms/wmt}) that allows users to build and run coupled Earth system models on
@@
 
+\begin{figure}
+  \caption{The \csdms{} Web Modeling Tool}
+  \begin{center}
+    \includegraphics[scale=.25]{wmt.eps}
+  \end{center}
+  \label{fig:wmt_screenshot}
+\end{figure}
 \begin{comment}
@@
 from the US National Science Foundation.
-
-\begin{figure}
-  \caption{The \csdms Web Modeling Tool}
-  \begin{center}
-    \includegraphics[scale=.25]{wmt.eps}
-  \end{center}
-  \label{fig:wmt_screenshot}
-\end{figure}
 

### 2014-07-14 paper.tex
Fixed all usages of \csdms{}.

@@
 
-\csdms, The Community Surface Dynamics Modeling System, is an NSF funded
+\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
 project whose focus is to aid a diverse community of earth and ocean system
@@
 basic smoke-tests of the models, (optionally) maintains the source code in a
-\csdms-hosted version control system, and provides easy downloads of the
+\csdms{}-hosted version control system, and provides easy downloads of the
 original source. The repository now contains over 200 models and tools (over
@@
 calling appication with the necessary level of control to connect components.
-\csdms, as well as other modeling frameworks such as
+\csdms{}, as well as other modeling frameworks such as
 ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
@@
 
-Another important goal of CSDMS is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales.
+Another important goal of \csdms{} is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales.
 Interdisciplinarity and coupling is a critical step towards advancing predictive modeling~\cite{voinov2010community}.
 
-CSDMS develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.
+\csdms{} develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.
 
@@
 clinics at the International Summer Institute at the University of Minnesota,
-and clinics at annual CSDMS Annual Meetings are the dominant forums for
+and clinics at annual \csdms{} Annual Meetings are the dominant forums for
 this testing and feedback.

### 2014-07-14 hutton_et_al_2014_wssspe.tex
Changed name of the tex file for the article.

--- 2014-07-14 paper.tex
@@
   Mark D. Piper$^{\dag}$,
-  Irina Overeem$^{\dag}$
-  Albert J. Kettner$^{\dag}$,
+  Irina Overeem$^{\dag}$,
+  Albert J. Kettner$^{\dag}$
 }
@@
 with the goal of building the set of tools necessary to model the
-earth system. \csdms{} is a place modelers look to for access to hundreds open
+earth system. Modelers look to \csdms{} for access to hundreds of open
 source surface-dynamics models and tools, as well as model metadata. Such a
-model repository increases model transparency and helps to eliminate the
-duplication of effors by presenting the current state of modeling efforts.
-To increase software sustainability, composability and interoperability \csdms{}
-promotes standards that define common modeling interfaces, semantic mediation
-between models, and model metadata. Through online resources and workshops,
-\csdms{} promotes software engineering best practices, which are foreign to many
-developers within our modeling community. For example, version control, unit
-testing, continious integration, test-driven development, well-written
-\emph{clean code} are all topics of the educational mission of \csdms{}.
+model repository increases model transparency and helps eliminate
+duplication by presenting the current state of modeling efforts.
+To increase software sustainability, composability and interoperability,
+\csdms{} promotes standards that define common modeling interfaces, semantic
+mediation between models, and model metadata. Through online resources and
+workshops, \csdms{} promotes software engineering best practices, which are
+unfamiliar to many developers within our modeling community. For example,
+version control, unit testing, continious integration, test-driven
+development, and well-written \emph{clean code} are all topics of the
+educational mission of \csdms{}.
 
@@
 service components.
-The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based client that provides a
+The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based application that provides a
 graphical interface to this framework and allows users to compose new models
@@
 \item Easily reuse and replicate research as software is directly locatable 
-\item Increase software visibility, \doi{} content is 5 times more likely to
+\item Increase software visibility: \doi{} content is 5 times more likely to
       deliver active links than content without.
-\item Provide e.g. funding agencies the ability to track usage, so measure
+\item Provide funding agencies with the ability to track usage, to measure
       impact.
@@
 
-The \bmi{} identifies the entry points into software components that provides a
-calling appication with the necessary level of control to connect components.
+The \bmi{} identifies the entry points into software components that provide a
+calling application with the necessary level of control to connect components.
 \csdms{}, as well as other modeling frameworks such as
@@
 
-%\item \emph{Maintenance}. Because modern browsers tend to adhere to web
-%      standards, which lead to fewer cross-compatibility issues than
-%      operating systems, only one version of WMT needs to be developed
-%      and maintained.
+\item \emph{Maintenance}. Because modern browsers tend to adhere to web
+      standards, which lead to fewer cross-compatibility issues than
+      operating systems, only one version of WMT needs to be developed
+      and maintained.
 
@@
 With \wmt{}, users can: \emph{run} single component models, build (2-way)
-\emph{coupled models}, view and \emph{edit} model parameters, \emph{share}
+\emph{coupled models}, view and \emph{edit} model parameters, and \emph{share}
 saved models with others in the community.
@@
 \end{figure}
+
 \begin{comment}
@@
 enrolled in intensive \csdms{}-run modeling courses. These courses
-provide an excellent use case for us to test usability and access to the
-tool. Students reported in post-course surveys that they were unfamiliar with
+provide an excellent use case for us to test usability and access to our
+tools. Students reported in post-course surveys that they were unfamiliar with
 the use of models on a high performance computing cluster beforehand (rated
@@
 
-\begin{comment}
-\huttonnote{Add something about model DOIs in the abstract}
-
-\huttonnote{
-\begin{itemize}
-\item improving the development process that leads to new software
-  \begin{itemize}
-  \item methods to develop sustainable software from the outset
-  \item effective approaches to reusable software created as a by-product of
-        research
-  \item impact of computer science research on the development of scientific software
-  \end{itemize}
-\end{itemize}
-}
-
-\huttonnote{This is Irina's section.}
-
-\huttonnote{
-\begin{itemize}
-\item building large and engaged user communities
-  \begin{itemize}
-  \item developing strong advocates
-  \item measurement of usage and impact
-  \end{itemize}
-\item encouraging industry's role in sustainability
-  \begin{itemize}
-  \item engagement of industry with volunteer communities
-  \item incentives for industry
-  \item incentives for community to contribute to industry-driven projects
-  \end{itemize}
-\item improving education and training
-  \begin{itemize}
-  \item best practices for providing graduate students and postdoctoral
-        researchers in domain communities with sufficient training in
-        software development
-  \item novel uses of sustainable software in education (K-20)
-  \item case studies from students on issues around software development in
-        the undergraduate or graduate curricula
-  \end{itemize}
-\end{itemize}
-}
-
-\huttonnote{This is Albert's section (he's rewriting this)}
-
-\huttonnote{
-\begin{itemize}
-\item recommending policy changes
-  \begin{itemize}
-  \item software credit, attribution, incentive, and reward
-  \item issues related to multiple organizations and multiple countries, such
-        as intellectual property, licensing, etc.
-  \item mechanisms and venues for publishing software, and the role of
-        publishers
-  \end{itemize}
-\end{itemize}
-}
-\end{comment}
-
 \end{document}

### 2014-07-14 hutton_et_al_2014_wssspe.tex
Added Albert's changes.

@@
   Irina Overeem$^{\dag}$,
-  Albert J. Kettner$^{\dag}$
+  Albert J. Kettner$^{\dag}$,
+  James P. M. Syvitski$^{\dag}$,
 }
@@
 with the goal of building the set of tools necessary to model the
-earth system. Modelers look to \csdms{} for access to hundreds of open
+earth system. Modelers use \csdms{} for access to hundreds of open
 source surface-dynamics models and tools, as well as model metadata. Such a
@@
 \csdms{} has implemented a
-Digital Object Identifier (\doi{}) system for software within the model
+Digital Object Identifier (\doi{}) system for software within its model
 repository to guarantee recognition to software contributors.
@@
 provides functionality to \emph{initialize}, \emph{update}, and
-\emph{finalize} a component. However, for more complicated (and 2-way)
-coupling components must implement the \bmi{} methods providing data access,
-and model metadata.
+\emph{finalize} a component. For more complicated (and 2-way)
+coupling, components must implement \bmi{} methods providing data access,
+and model metadata. For data access, \bmi{} defines a set of \emph{getter}
+and \emph{setter} methods that allow model components to present internal data
+for both reading and writing.
 
@@
 clinics at the International Summer Institute at the University of Minnesota,
-and clinics at annual \csdms{} Annual Meetings are the dominant forums for
+and clinics at \csdms{} Annual Meetings are the dominant forums for
 this testing and feedback.
@@
 the use of models on a high performance computing cluster beforehand (rated
-their familiarity as 1.5 average on a scale of 1-5), but ended the courses
-with familiarity and being more comfortable with running and changing
-simulations as well as visualizing results (average rating approaches 4 on
-a Likert scale of 1-5)~\cite{overeem2013strategies}.
+their familiarity as 1.5 average on a Likert scale of 1-5), but ended the
+courses with familiarity and being more comfortable with running and changing
+simulations as well as visualizing results (average rating approaches 4)
+~\cite{overeem2013strategies}.
 

### 2014-07-14 hutton_et_al_2014_wssspe.tex
Fixed spelling mistake.

@@
 unfamiliar to many developers within our modeling community. For example,
-version control, unit testing, continious integration, test-driven
+version control, unit testing, continuous integration, test-driven
 development, and well-written \emph{clean code} are all topics of the

### 2014-07-15 hutton_et_al_2014_wssspe.tex
Fixed spelling mistakes.

@@
 likely, is not paid to do this}), this mechanism must require minimal effort
-and the benifit to the developer should be clear.
+and the benefit to the developer should be clear.
 In the first few years of the \csdms{} project, we worked directly with model
@@
 original source. The repository now contains over 200 models and tools (over
-4M lines of code) and has recieved over 13k model downloads.
+4M lines of code) and has received over 13k model downloads.
 

### 2014-11-15 hutton_et_al_2014_wssspe.tex
Updated with reviewer comments.
Rewrote the Education section and changed a few items in the WTM
section. Most importantly, added Scott as author!

@@
   Mark D. Piper$^{\dag}$,
+  Scott D. Peckham$^{\dag}$,
   Irina Overeem$^{\dag}$,
@@
 
-\item \emph{Accessibility}. As a web-based application, if you have access
-      to the Internet, you have access to \wmt{}.
+\item \emph{Accessibility}. As a web-based application, users with internet
+      access, have access to \wmt{}.
 
@@
 
-\item \emph{Portability}. \wmt{} has a native JavaScript interface, so it can
-      be accessed on any modern web browser
+\item \emph{Portability}. \wmt{} has a native JavaScript interface.
+      JavaScript is an open standard and almost every computer is equipped
+      with a JavaScript-capable web browser.
 
-\item \emph{Maintenance}. Because modern browsers tend to adhere to web
-      standards, which lead to fewer cross-compatibility issues than
-      operating systems, only one version of WMT needs to be developed
-      and maintained.
+%\item \emph{Maintenance}. Because modern browsers tend to adhere to web
+%      standards, which lead to fewer cross-compatibility issues than
+%      operating systems, only one version of WMT needs to be developed
+%      and maintained.
 
@@
 
-\section{Education and Evaluation of \csdms{} Tools}
+\section{Education}
 \label{sec:education}
 
-%\wmt{} provides the ability to run models with an easy to use graphical
-%interface and intends to serve novice and more experienced modelers alike.
-%The WMT beta version was released in May 2014  as a sequel to the CMT.
+%\csdms{} constantly evaluates the effectiveness of its tools and standards
+%through user feedback from clinics and tutorials and one-on-one interviews
+%with \csdms{} members.  Modeling courses at the University of Colorado,
+%clinics at the International Summer Institute at the University of Minnesota,
+%and clinics at \csdms{} Annual Meetings are the dominant forums for
+%this testing and feedback.
 
-\csdms{} constantly evaluates the effectiveness of its tools and standards
-through user feedback from clinics and tutorials and one-on-one interviews
-with \csdms{} members.  Modeling courses at the University of Colorado,
-clinics at the International Summer Institute at the University of Minnesota,
-and clinics at \csdms{} Annual Meetings are the dominant forums for
-this testing and feedback.
+It has been our experience that new graduate students in the Earth sciences
+begin with little or no programming experience. However, their interest in
+programming is high and they see the importance of programming with regard
+to their thesis topics. More often than not, programming courses are not
+offered as part of their coursework and so, out of necessity to solve a
+particular problem, students end up teaching themselves how to program.
 
-Over the last 4 years over 150 graduate students - from geography, hydrology,
-geology, oceanography, earth sciences and engineering departments - have
-enrolled in intensive \csdms{}-run modeling courses. These courses
-provide an excellent use case for us to test usability and access to our
-tools. Students reported in post-course surveys that they were unfamiliar with
-the use of models on a high performance computing cluster beforehand (rated
-their familiarity as 1.5 average on a Likert scale of 1-5), but ended the
-courses with familiarity and being more comfortable with running and changing
-simulations as well as visualizing results (average rating approaches 4)
-~\cite{overeem2013strategies}.
+We believe providing these students with even a small base in programming and
+guidelines on best practices will go a long way. We have hosted several
+programming workshops, including one through Software Carpentry. The workshops
+were such a success that we are now planning an annual Software Carpentry
+bootcamp that is specifically targeted to incoming graduate students. The
+workshops fill up nearly immediately, which reflects a strong interest and
+need for basic software skills. We believe these workshops strike a good
+balance between teaching students about computer science topics and
+maintaining a pragmatic focus on problem solving.
 
-Course and clinic materials are posted online as educational resources.
-\csdms{} develops \wmt{}-specific help pages for all models currently
-available as components, so that users can easily learn about model
-parameters and input files. In addition, we are developing web-based video
-tutorials that show how to run models.
+%Over the last 4 years over 150 graduate students - from geography, hydrology,
+%geology, oceanography, earth sciences and engineering departments - have
+%enrolled in intensive \csdms{}-run modeling courses. These courses
+%provide an excellent use case for us to test usability and access to our
+%tools. Students reported in post-course surveys that they were unfamiliar with
+%the use of models on a high performance computing cluster beforehand (rated
+%their familiarity as 1.5 average on a Likert scale of 1-5), but ended the
+%courses with familiarity and being more comfortable with running and changing
+%simulations as well as visualizing results (average rating approaches 4)
+%~\cite{overeem2013strategies}.
+
+%Course and clinic materials are posted online as educational resources.
+%\csdms{} develops \wmt{}-specific help pages for all models currently
+%available as components, so that users can easily learn about model
+%parameters and input files. In addition, we are developing web-based video
+%tutorials that show how to run models.
 

### 2015-02-27 hutton_et_al_2014_wssspe.tex
Author additions for JSOR article.
The augmented WSSSPE2 article for the WSSSPE JSOR special issue.
These changes are mostly from Mark, Irina, and Albert.

@@
 earth-system models and tools to an international (67 country) community
-with the goal of building the set of tools necessary to model the
-earth system. Modelers use \csdms{} for access to hundreds of open
+with the goal of building the set of tools necessary to model
+earth surface processes. Modelers use \csdms{} for access to over two hundred open
 source surface-dynamics models and tools, as well as model metadata. Such a
 model repository increases model transparency and helps eliminate
-duplication by presenting the current state of modeling efforts.
+duplication by presenting the current state of community modeling efforts.
 To increase software sustainability, composability and interoperability,
@@
 workshops, \csdms{} promotes software engineering best practices, which are
-unfamiliar to many developers within our modeling community. For example,
+less familiar to many within our community. For example,
 version control, unit testing, continuous integration, test-driven
@@
 The mission of the Community Surface Dynamics Modeling System (\csdms{})
-~\cite{peckham2012component} is to help a diverse modeling community toward
+~\cite{peckham2012component} is to help move a diverse modeling community toward
 common goals and standards. This effort involves creating:
@@
   \item an \emph{efficient} and \emph{open} modeling community through
-        standards (Section ~\ref{sec:standards}) and education
-        (Section~\ref{sec:education})
+        standards (Section ~\ref{sec:standards}),
+        communication and education (Section~\ref{sec:education})
 \end{itemize}
@@
 service components.
-The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based application that provides a
-graphical interface to this framework and allows users to compose new models
-by connecting and configuring components in a simple, browser-based graphical
-interface.
+The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a
+web-based application that provides an interface to this
+framework.
+It allows users to compose new models by connecting and
+configuring components in a simple, browser-based graphical interface.
+Further, it provides a standardized interface to standalone models;
+once a user understands how to run a model in \wmt{},
+they can run any model in \wmt{}.
 
@@
 remains accessible beyond the lifetime of a specific research project. Code
-developers are required to distribute their codes under open-source licenses.
+developers are required to distribute their software under open-source licenses.
 
-\csdms{} has built an online model repository that not only contains model
-source code but also documents the submitted code as model metadata, runs
-basic smoke-tests of the models, (optionally) maintains the source code in a
-\csdms{}-hosted version control system, and provides easy downloads of the
-original source. The repository now contains over 200 models and tools (over
+\csdms{} has built an online model 
+repository\footnote{http://csdms.colorado.edu/wiki/Model\_download\_portal}
+(Figure~\ref{fig:model_repository_screenshot})
+where \csdms{} members can contribute code.
+The repository stores documentation on submitted code as model metadata,
+including a description, version number,
+a summary of inputs and outputs,
+and author contact information.
+Submitted code is given basic smoke tests,
+then it is placed under version control in a 
+GitHub repository\footnote{https://github.com/csdms-contrib}
+(if the contributor already maintains the code on GitHub, 
+the code is forked from their repository), 
+where it can be forked or cloned by other users.
+The repository currently contains over 200 models and tools (over
 4M lines of code) and has received over 13k model downloads.
+
+\begin{figure}
+  \caption{The \csdms{} Model Repository}
+  \begin{center}
+    \includegraphics[scale=.25]{csdms_model_repository.eps}
+  \end{center}
+  \label{fig:model_repository_screenshot}
+\end{figure}
 
@@
 Only stable versions of open source
-software that are physically hosted through the \csdms{} repository will be
+software that are physically hosted through the \csdms{} model repository are
 assigned a \doi{}. For model software citations, \csdms{} recommends the
@@
 (Figure~\ref{fig:wmt_screenshot}) and a RESTful
-server-side database and API (the \wmt{} server\footnote{https://github.com/csdms/wmt}) that allows users to build and run coupled Earth system models on
+server-side database and API (the \wmt{} server\footnote{https://github.com/csdms/wmt})
+that allows users to build and run coupled models of earth surface processes on
 a high-performance computing cluster from a web browser. \wmt{} was designed
-with four objectives:
+with three objectives:
 \begin{itemize}
 
-\item \emph{Accessibility}. As a web-based application, users with internet
-      access, have access to \wmt{}.
+\item \emph{Accessibility}. As a web-based application, access
+      to the Internet implies access to \wmt{}.
 
 \item \emph{Integration}. Easily hyperlink from \wmt{} to resources on the 
-      \csdms{} modeling portal (model documentation, labs, lectures,
+      \csdms{} portal (model documentation, labs, lectures,
       tutorials)
@@
 
-\item \emph{Portability}. \wmt{} has a native JavaScript interface.
-      JavaScript is an open standard and almost every computer is equipped
-      with a JavaScript-capable web browser.
-
-%\item \emph{Maintenance}. Because modern browsers tend to adhere to web
-%      standards, which lead to fewer cross-compatibility issues than
-%      operating systems, only one version of WMT needs to be developed
-%      and maintained.
+\item \emph{Portability}. \wmt{} has a native JavaScript interface. JavaScript
+      is an open standard and almost every computer is equipped with a
+      JavaScript-capable web browser.
 
 \end{itemize}
+
+\begin{figure}
+  \caption{The \csdms{} Web Modeling Tool}
+  \begin{center}
+    \includegraphics[scale=.25]{wmt.eps}
+  \end{center}
+  \label{fig:wmt_screenshot}
+\end{figure}
+
 With \wmt{}, users can: \emph{run} single component models, build (2-way)
@@
 saved models with others in the community.
+
+%% \wmt{}, and the software stack on which it is built, are portable.
 
@@
 
-\begin{figure}
-  \caption{The \csdms{} Web Modeling Tool}
-  \begin{center}
-    \includegraphics[scale=.25]{wmt.eps}
-  \end{center}
-  \label{fig:wmt_screenshot}
-\end{figure}
-
 \begin{comment}
@@
 
-\section{Education}
+\section{Communication and Education}
 \label{sec:education}
 
-%\csdms{} constantly evaluates the effectiveness of its tools and standards
-%through user feedback from clinics and tutorials and one-on-one interviews
-%with \csdms{} members.  Modeling courses at the University of Colorado,
-%clinics at the International Summer Institute at the University of Minnesota,
-%and clinics at \csdms{} Annual Meetings are the dominant forums for
-%this testing and feedback.
+%\wmt{} provides the ability to run models with an easy to use graphical
+%interface and intends to serve novice and more experienced modelers alike.
+%The WMT beta version was released in May 2014  as a sequel to the CMT.
 
-It has been our experience that new graduate students in the Earth sciences
-begin with little or no programming experience. However, their interest in
-programming is high and they see the importance of programming with regard
-to their thesis topics. More often than not, programming courses are not
-offered as part of their coursework and so, out of necessity to solve a
-particular problem, students end up teaching themselves how to program.
+\subsection{The CSDMS Portal}
 
-We believe providing these students with even a small base in programming and
-guidelines on best practices will go a long way. We have hosted several
-programming workshops, including one through Software Carpentry. The workshops
-were such a success that we are now planning an annual Software Carpentry
-bootcamp that is specifically targeted to incoming graduate students. The
-workshops fill up nearly immediately, which reflects a strong interest and
-need for basic software skills. We believe these workshops strike a good
-balance between teaching students about computer science topics and
+The \csdms{} portal\footnote{http://csdms.colorado.edu}
+is the information clearinghouse for the community.
+Through the portal,
+members can get the latest news,
+access model metadata and source code (Section~\ref{sec:organizing}),
+view conference and workshop announcements,
+view recent job postings,
+and access \csdms{} reports, presentations, and publications.
+%% including educational presentations and short course
+%% materials from \csdms{}-hosted or sponsored workshops.
+The portal uses the open-source MediaWiki software,
+so \csdms{} members can edit, update and extend existing material,
+as well as contribute new material to the portal.
+
+\subsection{User Training}
+
+\csdms{} provides user training in modeling and model coupling
+on high-performance computing clusters
+through short courses held at 
+the University of Colorado,
+the International Summer Institute at the University of Minnesota,
+and at \csdms{} Annual Meetings.
+Course materials are posted to the \csdms{} portal as educational resources.
+In addition to training community members,
+the clinics also provide an excellent opportunity 
+for feedback on using \csdms{} software tools.
+
+\subsection{Programming Courses}
+
+It has been our experience that new graduate students in the Earth
+sciences begin with little or no programming experience. However,
+their interest in programming is high and they see the importance of
+programming with regard to their thesis topics. More often than not,
+programming courses are not offered as part of their coursework and
+so, out of necessity to solve a particular problem, students end up
+teaching themselves how to program.
+
+We believe providing these students with even a small base in programming 
+and guidelines on best practices will go a long way. We have
+hosted several programming workshops, including one through Software
+Carpentry. The workshops were such a success that we are now planning
+an annual Software Carpentry bootcamp that is specifically targeted
+to incoming graduate students. The workshops fill up nearly
+immediately, which reflects a strong interest and need for basic
+software skills. We believe these workshops strike a good balance
+between teaching students about computer science topics and
 maintaining a pragmatic focus on problem solving.
-
-%Over the last 4 years over 150 graduate students - from geography, hydrology,
-%geology, oceanography, earth sciences and engineering departments - have
-%enrolled in intensive \csdms{}-run modeling courses. These courses
-%provide an excellent use case for us to test usability and access to our
-%tools. Students reported in post-course surveys that they were unfamiliar with
-%the use of models on a high performance computing cluster beforehand (rated
-%their familiarity as 1.5 average on a Likert scale of 1-5), but ended the
-%courses with familiarity and being more comfortable with running and changing
-%simulations as well as visualizing results (average rating approaches 4)
-%~\cite{overeem2013strategies}.
-
-%Course and clinic materials are posted online as educational resources.
-%\csdms{} develops \wmt{}-specific help pages for all models currently
-%available as components, so that users can easily learn about model
-%parameters and input files. In addition, we are developing web-based video
-%tutorials that show how to run models.
 
@@
 representatives in addition to members from academic institutions. The
-initiative has been growing with more than 150 people per year. \csdms{} now
-fosters a community of approximately 1200 scientists who work on prediction
-of the movement of fluids, and the production, erosion, transport, and
+initiative has been growing, adding more than 150 people per year. \csdms{} now
+fosters a community of over 1200 scientists who work on determining
+the movement of fluids, and the production, erosion, transport, and
 deposition of sediment and nutrients in landscapes and seascapes.

### 2015-03-02 article.tex
Renamed tex and bib files as "article".
Removed abstract and authors for JORS submission.

--- 2015-02-27 hutton_et_al_2014_wssspe.tex
@@
 \usepackage[all]{hypcap}
+\usepackage{doi}
 \urlstyle{rm}
@@
 
-\author{
-  Eric W. H. Hutton$^{\dag}$,
-  Mark D. Piper$^{\dag}$,
-  Scott D. Peckham$^{\dag}$,
-  Irina Overeem$^{\dag}$,
-  Albert J. Kettner$^{\dag}$,
-  James P. M. Syvitski$^{\dag}$,
-}
+%\author{
+%  Eric W. H. Hutton$^{\dag}$,
+%  Mark D. Piper$^{\dag}$,
+%  Scott D. Peckham$^{\dag}$,
+%  Irina Overeem$^{\dag}$,
+%  Albert J. Kettner$^{\dag}$,
+%  James P. M. Syvitski$^{\dag}$,
+%}
 
@@
 
-\begin{abstract}
+%\begin{abstract}
 
-\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
-project whose focus is to aid a diverse community of earth and ocean system
-model \emph{users} and \emph{developers} to use and create robust software
-quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
-earth-system models and tools to an international (67 country) community
-with the goal of building the set of tools necessary to model
-earth surface processes. Modelers use \csdms{} for access to over two hundred open
-source surface-dynamics models and tools, as well as model metadata. Such a
-model repository increases model transparency and helps eliminate
-duplication by presenting the current state of community modeling efforts.
-To increase software sustainability, composability and interoperability,
-\csdms{} promotes standards that define common modeling interfaces, semantic
-mediation between models, and model metadata. Through online resources and
-workshops, \csdms{} promotes software engineering best practices, which are
-less familiar to many within our community. For example,
-version control, unit testing, continuous integration, test-driven
-development, and well-written \emph{clean code} are all topics of the
-educational mission of \csdms{}.
+%\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
+%project whose focus is to aid a diverse community of earth and ocean system
+%model \emph{users} and \emph{developers} to use and create robust software
+%quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
+%earth-system models and tools to an international (67 country) community
+%with the goal of building the set of tools necessary to model
+%earth surface processes. Modelers use \csdms{} for access to over two hundred open
+%source surface-dynamics models and tools, as well as model metadata. Such a
+%model repository increases model transparency and helps eliminate
+%duplication by presenting the current state of community modeling efforts.
+%To increase software sustainability, composability and interoperability,
+%\csdms{} promotes standards that define common modeling interfaces, semantic
+%mediation between models, and model metadata. Through online resources and
+%workshops, \csdms{} promotes software engineering best practices, which are
+%less familiar to many within our community. For example,
+%version control, unit testing, continuous integration, test-driven
+%development, and well-written \emph{clean code} are all topics of the
+%educational mission of \csdms{}.
 
-\end{abstract}
+%\end{abstract}
 
@@
 \footnote{https://github.com/csdms/standard\_names}
-are a common language for variable names exchanged between models. They play
+are a common language for variable names exchanged between models
+~\cite{peckham2014csdms}. They play
 an important role in the \bmi{} as they provide a mapping of a model's
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{plainurl}
 
-\bibliography{wssspe-citations}{}
+\bibliography{article-citations}{}
 
